### 2023-05-21 mlm-lmc.Rmd
Create mlm-lmc.Rmd

---
title: "latent-mean-center-mlm"
author: "Ethan M. McCormick"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#| warning: false
#| message: false
#| include: false

packages <- c("dplyr", "lmerTest", "brms")
invisible(lapply(packages, library, character.only = TRUE))
if (require(cmdstanr)) {
  options(brms.backend = "cmdstanr")
}
```

```{r}
btDat <- read.csv("~/Dropbox/Professional/projects/longitudinal-primer/codebook/data/feedback-learning.csv") %>%
  mutate(wave = wave - 2,
         age = (age - mean(age, na.rm = TRUE)) / sd(age, na.rm = TRUE),
         age2 = age^2) %>%
  group_by(id) %>%
  mutate(modul_id_mcent = modularity - mean(modularity, na.rm = TRUE),
         modul_id_mean = mean(modularity, na.rm = TRUE))
```

```{r}
lmer.fit <- lmer(learning.rate ~ age + age2 + 
                   modul_id_mcent + modul_id_mean + (1 | id),
                 data = btDat)

summary(lmer.fit)
```

```{r}
model <- bf(
  learning.rate ~ 1 + age + age2 + modul_id_mcent + modul_id_mean + (1 | id),
  #age + age2 + beta + beta2 ~ 1,
  nl = F
) +
  gaussian()

p <- get_prior(model, data = btDat) %>%
  mutate(
    prior = case_when(
      class == "b" & coef == "Intercept" ~ "normal(0, 1.5)",
      class == "sd" & coef == "Intercept" ~ "student_t(7, 0, 1)",
      TRUE ~ prior
    )
  )

job::job({fit1 <- brm(
  model,
  data = btDat,
  prior = p,
  cores = round(parallel::detectCores(logical = FALSE) %/% 2),
  iter = 10000,
  control = list(adapt_delta = 0.99),
  file = "models/braintime-fixed"
)})
summary(fit1)
```

```{r}
model <- bf(
  learning.rate ~ alpha + b1*age + b2*age2 + b3*(modularity - modC),
  alpha ~ 1 + (1 |id),
  b1 + b2 + b3 ~ 1,
  nlf(modC ~ modCi),
  modCi ~ 1 + (1 | id),
  nl = TRUE
) +
  gaussian()

p <- get_prior(model, data = btDat) %>%
  mutate(
    prior = case_when(
      class == "b" & coef == "Intercept" ~ "normal(0, 1.5)",
      class == "sd" & coef == "Intercept" ~ "student_t(7, 0, 1)",
      TRUE ~ prior
    )
  )

job::job({fit2 <- brm(
  model,
  data = btDat,
  prior = p,
  cores = round(parallel::detectCores(logical = FALSE) %/% 2),
  iter = 10000,
  control = list(adapt_delta = 0.99),
  file = "models/braintime-lmc"
)})
summary(fit2)
```