### 2023-10-31 simulation.qmd
Wrangle data for brms

---
title: "Simulation"
execute: 
  cache: true
format: 
  html:
    toc: true
    code-fold: true
---

```{r}
#| label: options

library(posterior)
library(mvtnorm)
library(cmdstanr)
library(ggdist)
library(brms)
library(tidyverse)

N_CORES <- min(parallel::detectCores(), 8)
options(
  mc.cores = N_CORES,
  brms.backend = if(require(cmdstanr)) "cmdstanr",
  brms.threads = N_CORES %/% 4
)
```

In this document we specify, simulate, and validate a univariate latent means brms model. Code based on Joran's blog post at <https://experienced-sampler.netlify.app/post/stan-hierarchical-ar/>.

# Data

We first set up the data simulation. I used [this code](https://experienced-sampler.netlify.app/post/stan-hierarchical-ar/#testing) but put it inside a function.

```{r}
#' Simulate ar(1) data
#' 
#' Adapted from <https://experienced-sampler.netlify.app/post/stan-hierarchical-ar/#testing>
#'
#' @param I number of individuals
#' @param t number of time points per individual
#' @param sigma residual standard deviation of observations
#' @param alpha_hat intercept mean
#' @param alpha_scale intercept standard deviation
#' @param beta_hat ar(1) slope mean
#' @param beta_scale ar(1) slope standard deviation
#'
#' @return a list of variables to send to Stan
#'
#' @examples simulate()
simulate <- function(
    I = 50, 
    t = 25, 
    sigma = 1, 
    alpha_hat = 4, 
    alpha_scale = 1, 
    beta_hat = 0.4, 
    beta_scale = 0.1
) {

  # Define data
  N <- I*t # total sample size
  individual <- rep(1:I, each = t)
  time <- rep(1:t, I)
  alphas <- rnorm(I, alpha_hat, alpha_scale) # person-specific intercepts
  betas <- rnorm(I, beta_hat, beta_scale) # person-specific slopes
  for (i in 1:I) {
    # The while loop avoids non-stationary AR processes
    # See Hamilton  pg. 259
    while(betas[i] <= -1 | betas[i] >= 1) {
      betas[i] <- rnorm(1, beta_hat, beta_scale)
    }
  }

  # Determine first observations for everyone. The variance for this first 
  # observation is different than for the subsequent ones and so it needs to be 
  # samples separately
  IndT1 <- match(unique(individual), individual)

  # Determine variance at first measurement for everyone 
  # (depends on their AR-parameter)
  sigmaT1 <- rep(NA, I)

  for (k in 1:I) {
    sigmaT1[k] <- sigma/(1-((betas[k])^2))
  }

  # First create storage matrices for non-centered and centered y-scores.
  # We need centered values, because of we use person-centered values as predictors, 
  # alpha will be equal to individual means instead of individual intercepts 
  # which are less informative.
  Y <- rep(NA, N)
  Yc <- rep(NA, N)

  # Draw first observation for each individual first
  for (l in 1:I) {
    Y[IndT1[l]] <- rnorm(1, alphas[l], sigmaT1[l])
    Yc[IndT1[l]] <-  Y[IndT1[l]] - alphas[l]
  }

  # Draw subsequent observations
  for (m in 1:N) {
    # This if statement makes sure I don't try to predict a persons first 
    # observation which is impossible there is no measurement before the first 
    # observation and so no predictor values for that observation
    if (time[m]>1) {
      Y[m] <- rnorm(1, (alphas[individual[m]] + betas[individual[m]]*Yc[m-1]), sigma)
      Yc[m] <-  Y[m] - alphas[individual[m]]
    }
  }

  return(
    list(
      I = I,
      T = t,
      N = N,
      sigma = sigma, 
      alpha_hat = alpha_hat, 
      alpha_scale = alpha_scale, 
      beta_hat = beta_hat, 
      beta_scale = beta_scale,
      individual = individual,
      time = time,
      y = Y
    )
  )

}
```

The function returns a list that we can pass to Stan. It also contains the data generation parameters.

```{r}
set.seed(31121)
dat <- simulate()
glimpse(simulate())
```

## Stan model

Let us then compile and sample from the cmdstanr model

```{r}
# model <- cmdstan_model("model-ar1.stan", compile = FALSE)
# model$format(
#   canonicalize = list("deprecations"), 
#   overwrite_file = TRUE, 
#   backup = FALSE
# )
# model$compile()
model <- cmdstan_model(exe_file = "model-ar1")
path <- "models/cmdstanr-example.rds"
if (!file.exists(path)) {
post <- model$sample(
  data = dat, 
  adapt_delta = 0.99
) |> 
  as_draws_df()
saveRDS(post, path)
} else {post <- readRDS(path)}
```

Visualize estimated and defined parameters

```{r}
post <- post |> 
  select(starts_with("beta_"), ends_with("_raw")) |> 
  rename_with(~str_remove(., "_raw"))
post |> 
  pivot_longer(everything()) |> 
  ggplot(aes(value, name)) +
  scale_x_continuous(
    "Parameter value"
  ) +
  stat_histinterval(
    normalize = "xy",
    scale = 0.25, 
    size = 1
  ) +
  geom_point(
    data = as_tibble(dat) |> 
      select(sigma:beta_scale) |> 
      distinct() |> 
      pivot_longer(everything()),
    position = position_nudge(y = -0.1),
    color = "red"
  ) +
  theme(
    axis.title.y = element_blank()
  )
```

## brms

```{r}
dat <- as_tibble(dat)
dat <- dat |> 
  mutate(yt1 = lag(y), .by = individual)
dat
```