### 2023-12-04 data.qmd

---
title: Example datasets
---

Here we download and wrangle the example datasets used in our analyses.

```{r}
#| label: setup

library(tidyverse)
```

Throughout, we use an example dataset from "A primer on two-level dynamic structural equation models for intensive longitudinal data in Mplus" [@mcneishPrimerTwolevelDynamic2020]. Those data are available on the OSF (<https://osf.io/wuprx/>). Here we download, clean, and save the data for use in our analyses.

```{r}
#| label: mh-download

path <- "data/mh.zip"
url <- "https://files.osf.io/v1/resources/wuprx/providers/osfstorage/5bfc839601593f0016774697/?zip="
if (!file.exists(path)) {
  dir.create(dirname(path), showWarnings = FALSE)
  download.file(
    url,
    destfile = path
  )
  unzip(path, exdir = "data/mh/")
}
```

That extracted all the data and Mplus files to `data/` in the project root. We then wrangle the main data file and save for our analyses:

```{r}
#| label: mh-wrangle

d <- read_csv(
  "data/mh/Data/Two-Level Data.csv", 
  col_names = c("urge", "dep", "js", "hs", "person", "time")
) %>% 
  select(
    person, time, js, hs, urge, dep
  ) |> 
  mutate(
    person = factor(person),
    time = as.integer(time-1)
  )
write_rds(d, "data/mcneish-hamaker.rds")
```