### 2023-12-04 index.qmd

---
author:
  - name: Matti Vuorre
    url: https://vuorre.com
    orcid: 0000-0001-5052-066X
    email: mjvuorre@uvt.nl
    affiliation:
      - ref: 1
  - name: Ethan McCormick
    affiliation:
      - ref: 3
  - name: Mauricio Garnier-Villarreal
    affiliation:
      - ref: 4
  - name: Joran Jongerling
    orcid: 0000-0001-5697-1381
    affiliation:
      - ref: 2
affiliations:
  - id: 1
    name: Tilburg University
  - id: 2
    name: Tilburg University
  - id: 3
    name: todo
  - id: 4
    name: todo
---

{{< include README.qmd >}}