### 2024-02-19 README.qmd
Update readme

---
title: Introduction to latent means models in R
format: gfm
---

Repository for our manuscript discussing an [R](https://cran.r-project.org/) and [brms](https://cran.r-project.org/web/packages/brms/) implementation of latent mean centering in multilevel models of longitudinal psychological data.

- Repo on GitHub: <https://github.com/mvuorre/latent-mean-centering-ms>
- Archived repo: tbd
- Online: tbd
- preprint: tbd

## Reproduce

Everything is coded in [R](https://cran.r-project.org/) and bunched together as a [Quarto](https://quarto.org/) website. To reproduce, clone the repo and render the Quarto documents:

```bash
# Download materials
git clone https://github.com/mvuorre/latent-mean-centering-ms.git
cd latent-mean-centering-ms

# Set up R environment
Rscript -e 'renv::restore(prompt = FALSE)'

# Modify environment variables in .Renviron.example & rename it to .Renviron

# Run code and render outputs
quarto render
```

Once the computations are finished (this will take a long time depending on your computer) the rendered documents are in `docs/`.

## Contribute

Contributions preferably via pull requests. You can also grab the rendered manuscript (Word, PDF) from the latest [release](https://github.com/mvuorre/latent-mean-centering-ms/releases) for comments & edits.